Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-07-28 paper-amia-summit-2018/manuscript.tex
update discussion section

\section*{Discussion}
This study is the first effort to evaluate the automatic segmentation of e-Coaching emails. Experimental results indicate that SVM is the best model among all machine learning methods considered for this study. SVM achieved 0.990 F1-measure in overall, 0.845 and 0.995 F1-measures for detecting ``new segment'' and ``same segment'', respectively. The robust performance of SVM provides the evidence that machine learning models are capable to learn conceptual information from clinical exchange. It also indicates that topic features are more important than lexical features, even when deep learning methods are employed. Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and it can be successfully applied to other domains, in which discourse segmentation is a preliminary step for annotation.

Punctuation mark and topic model-based features made a significant improvement in performance of all machine learning methods. Among all topic model-based feature, the DL-LDA-based topic feature provides the highest performance in ``new segment'' and ``same segment'' classification as well as weighted average over ``new segment'' and ``same segment'' classification results. In nearly all cases, ML methods perform better, when lexical features are used in combination with punctuation and topic model-based features. This results also indicate that segmentation performance might be improved by adding additional relevant features.

@@

In this paper, results are reported by each class to avoid confusion regarding overall model performance due to severe class imbalance. In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets. AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the ROC curves in Figure~\ref{fig:roc-curves}. Although NB provides lowest classification results, it shows moderate AUC values, achieving 0.978 AUC when only lexical features are used and 0.977 AUC when a combination of lexical, punctuation and topic features are used. On the other hand, KNN and SVM exhibit 0.972 and 0.986 AUCs when only lexical features are used; and 0.966 and 0.980 AUCs when a combination of lexical, punctuation and topic model-based features are used. Finally, LSTM demonstrates the highest AUC value among all machine learning models, achieving 0.996 AUC, while GRU achieves only 0.2\% lower AUC than LSTM. \\

@@

We observed moderate performance of RNN, in particular, LSTM and GRU for the text segmentation. We believe that RNN will perform better if a larger dataset is utilized. In this study, we employed 3,138 examples of boundary case which limit to achieve the best performance. We also observed that GRU performs better than LSTM, which was observed in previous studies\cite{chung2014empirical}.

Although punctuation mark plays an important role in segmentation boundary detection and large numbers of errors were encountered by the false positive of boundary identification. For example, a text segment from an e-Coaching email \textit{``A typical day in regards to fruit and vegetable has me eating about a serving at breakfast (our caf has cut up fruit) and then maybe a piece of fruit later in the day or as a snack. Vegetable tends to be a side serving at lunch and dinner and I get celery or carrot cuts with dressing for a snack a lot of times. I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I quotemark m hungry rather than the junk food I tend to eat.''} incorrectly segmented after the first sentence where a punctuation mark was encountered. Similarly, additional information is a common cause for misclassification of an email segment into multiple segments. For instance, the first sentence of the above email segment represents a positive commitment to adolescent's behavior change then next two sentences provide an additional information to support their commitment.

Our proposed approach is novel for the segmentation of e-Coaching emails since previous studies mainly focused on the segmentation of text into sections, headers and sentences. However, this study segmented clinical exchange into groups of MI behaviors which will significantly reduce the amount of resource and time required to segment clinical exchange manually. Furthermore, these segmentation models could be integrated with auto coding classifiers to build a software pipeline for automated annotation of exchanges.

>>>>

\section*{Discussion}
This study is the first effort to evaluate the automatic segmentation of e-Coaching emails. Experimental results indicate that CRNN is the best model among all machine learning methods considered for this study. CRNN achieved 0.989 F1-measure in overall and 0.825 F1-measure for detecting ``new segment''. The robust performance of CRNN provides the evidence that machine learning models are capable to learn conceptual information from the clinical exchange. It also indicates that punctuation and POS features are more important than word embeddings or lexical features, even when deep learning methods are employed. Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and it can be successfully applied to other domains, in which discourse segmentation is a preliminary step for annotation.

Punctuation mark and POS features made a significant improvement in the performance of machine learning and all deep learning methods. In all cases, every method performs better, when word embeddings or lexical features are used in combination with punctuation and POS features. This results also indicate that segmentation performance might be improved by adding additional relevant features. \\

@@

In this paper, results are reported by ``new segment'' class as well as weighted average over both classes to avoid confusion regarding overall model performance due to severe class imbalance. In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets. AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the Table~\ref{tab:result_roc} and ROC curves in Figure~\ref{fig:roc-curves}. Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used. Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively. \\

@@

Although punctuation mark plays an important role in segmentation boundary detection, large numbers of errors were encountered by the false positive of boundary identification. For example, a text segment from an e-Coaching email \textit{``A typical day in regards to fruit and vegetable has me eating about a serving at breakfast (our caf has cut up fruit) and then maybe a piece of fruit later in the day or as a snack. Vegetable tends to be a side serving at lunch and dinner and I get celery or carrot cuts with dressing for a snack a lot of times. I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I quotemark m hungry rather than the junk food I tend to eat.''} incorrectly segmented after the first sentence where a punctuation mark was encountered. Similarly, additional information is a common cause for misclassification of an email segment into multiple segments. For instance, the first sentence of the above email segment represents a positive commitment to adolescent's behavior change then next two sentences provide an additional information to support their commitment.

Our proposed approach is novel for the segmentation of e-Coaching emails since previous studies mainly focused on the segmentation of text into sections, headers and sentences. However, this study segmented clinical exchange into groups of MI behaviors which will significantly reduce the amount of resource and time required to segment clinical exchange manually. Furthermore, these segmentation models could be integrated with auto coding classifiers to build a software pipeline for automated annotation of clinical exchanges.
